Exploring Potential 6LoWPAN Traffic Side Channels

### 2017-04-07 Writing/introduction.tex
Update
River publisher ready version 1.0.

\section{Introduction}
\textit{
\begin{itemize}
	\item Security options in 6LoWPAN: 802.15.4 Security and DTLS.
\end{itemize}
}

>>>>

\subsection*{Abstract}
The Internet of Things (IoT) has become a reality: small connected devices feature in everyday objects
including childrens' toys, TVs, fridges, heating control units, etc. Supply chains feature sensors
throughout, and significant investments go into researching next-generation healthcare, 
where sensors monitor wellbeing. A future in which sensors and other (small) devices interact to create sophisticated applications seems just around the corner. All of these applications have a fundamental need for security and privacy and thus cryptography is deployed as part of an attempt to secure them. In this paper we explore a particular type of flaw, namely side channel information, on the protocol level that can exist despite the use of cryptography. Our research investigates the potential for utilising packet length and timing information (both are easily obtained) to extract interesting information from a system. We find that using these side channels we can distinguish between devices, different programs running on the same device including which sensor is accessed. We also find it is possible to distinguish between different types of ICMP messages despite the use of encryption. Based on our findings, we provide a set of recommendations to efficiently mitigate these side channels in the IoT context.

%\keywords{6LoWPAN, Side Channels, Traffic Analysis}

\section{Introduction}
The expression `Internet of Things' (IoT) can refer to a multitude of objects and protocols, which share that they have been purposefully designed for resource constraint environments. Whereas the typical TCP/IP network stack produces considerable overhead to achieve quality of service for applications that are based on it, the nature of many IoT `things' is such that a full implementation of it would not be practical. Often `things' are sensor, which are devices that have to function on little resources (most importantly power). Thus a whole host of new networking protocols have been developed over the years to cater for such resource constrained devices: 6LoWPAN is the `tiny' version of IPv6, UDP tends to be used instead of TCP/IP, DTLS can be used for end-to-end security or one can directly invoke 802.15.4 security which is part of 6LoWPAN, and finally CoAP(s) is the replacement for HTTP(s). Thus there are two options (802.15.4, and DTLS) to secure communications between the `things' and a server/gateway.

Implementing cryptography correctly and securely has proven to be a massive challenge as evidenced by the multitude of implementation attacks over the years. Triggered off by research that showed how to utilise additional information via timing and power side channels\cite{DPA}, many different flavours of side channel attacks were discovered over the last decade. Many attacks use phyiscal information (such as low level execution timings or power consumption) to recover secret keys, but many other attacks use protocol level information (such as packet lengths, types of packets or protocol messages) to recover information about plaintexts, devices in the network, or the network itself. 
There exists a considerable body of work in the context of conventional, i.e. HTTPs over TCP/IP network, but the applicability of (some) of these attacks in the context of a typical IoT protocol stack is lacking. This is the gap that we would like to address with this work.

This paper is structured as follows: after reviewing some relevant attack paths for HTTPs over TCP/IP in the following subsection, we provide a brief introduction to the necessary protocol and network features in \Cref{sec: IotProtocols}.  We discuss the impact of packet length leakage in \Cref{sec: PacketLen}, followed by an analysis of the response time leakage in \Cref{sec: ResponseTime}. We summarise our work in \Cref{sec: Conclusion}.

\subsection{Related Work\label{subsec: RelatedWork}}
%Write about attacks that exploit packet length and response times in the context of TCP/IP. Reflect on web applications and how the same might happen if more interesting applications might be deployed that involve IoT objects.

%Literatures about Traffic Analysis...
Traffic Analysis is well studied in the context of encrypted Internet traffic, especially for web applications based on HTTPs and TCP/IP. The landmark study by Chen et al. \cite{WebSidechannel} discussed different side channel attacks against web applications and \cite{SuggestBox} studied the practicability of an attack specifically targeted Google and Bing search boxes. Later work by Mather and Oswald \cite{PinpointWeb} proposed the use of Mutual Information to pinpoint the potential leakage points in web traffic. For non-HTTPs applications, the papers \cite{AppleMsg}, \cite{Language} and \cite{VideoTraffic} described attacks against encrypted text, voice and video traffic respectively. Machine learning is widely used to analyse the traffic, and behaviours of different classifiers are studied by \cite{HClassifier} and \cite{Peekaboo}. Based on all these published works we can conclude that two features, the packet length and response time, are the most exploited ones among all attacks. Different countermeasures were studied by \cite{TrafficMorphing}, \cite{HTTPOS} and \cite{FTE}.

Reflecting on IoT applications, we stipulate that most of these attacks may still be applicable, as we intend to demonstrate in this paper. Considering the future vision that IoT devices could be indeed connected to the Internet with even more sensitive data flowing over different networks, the task of designing secure IoT applications becomes increasingly challenging.

%Literatures about 6LoWPAN security
With regard to the aspect of protocol design, the recent paper \cite{6LoWPANAtk} summarised some known flaws of 6LoWPAN, including its susceptibility to the Fragmentation Attack\cite{FragAtk}, Sinkhole Attack\cite{Sinkhole}, Hello Flood Attack\cite{HelloFlood}, Wormhole Attack\cite{Wormhole} and Blackhole Attack\cite{Blackhole}. In addition, \cite{802154SecIssues} reported certain problematic designs in 802.15.4 security\cite{802154}. However we do not discuss further these particular design flaws as they touch on a different aspect of the security issues in 6LoWPAN compared to what we address in this paper.

### 2017-04-07 Writing/iotprotocols.tex
Finalised V0.9.
Final draft 0.99.
River publisher ready version 1.0.

\section{A Typical IoT Protocol Stack}
%A typical sensor network is built as described in \ref{Protocols}.
%IoT Protocols
Many protocols are proposed for different IoT applications adapting to various requirements. For example, some smart houses simply use WiFi for connectivity and VANETs\footnote{Vehicular ad hoc networks} may adopt DSRC\cite{DSRC}.

%WSN Protocols
In this paper we focused on 6LoWPAN\cite{rfc4944} which is based on 802.15.4\cite{802154}. These standards are designated for constrained environments such as Wireless Sensor Networks,  but other competing standards exist at different layers. Bluetooth Low Energy(BLE)\cite{BLE} is a strong competitor to 802.15.4 as well as the LiFi\cite{LiFi} technology. Zigbee\cite{Zigbee} is originally intended as a collective protocol over 802.15.4 but it has been recently adapted to IP-based network in ZigbeeIP\cite{ZigbeeIp} so to compete with 6LoWPAN. The RIME stack\cite{RIME} proposed a set of non-layered primitives over 802.15.4 but it is likely to be phased-out probably due to the reluctance of interoperability with the TCP/IP protocol stack of Internet.

%6LoWPAN OS
6LoWPAN is supported by several competing IoT Operating Systems, including Contiki OS\cite{Contiki}, OpenWSN\cite{OpenWSN}, FreeRTOS\cite{FreeRTOS} and the recent RIOT\cite{RIOT}. Contiki OS is chosen for our experiments for its customisability.

\subsection{Our experimental network}
%Say something about devices, and choice of OS, etc.
%Hardware
Our experiment networks are constructed using devices constituted of TelosB\cite{TelosB} and CC2538\cite{CC2538}. TelosB is a low cost sensor powered by MSP430 with AES co-processor representing the low-end devices. CC2538 is the high end device powered by ARM Cortex-M3 with multiple cryptographic processors including AES, RSA, SHA-2 and ECC, suggesting it for security usage.
%Should I cite our previous paper...?

%Software
On the software side, we by principle adopted the default settings of Contiki OS, except enabling 802.15.4 Security\cite{802154} upon its study. Note that Contiki MAC\cite{ContikiMAC} is chosen by default over TSCH\cite{TSCH}, albeit we would argue the channel information of TSCH could be an additional leakage source. For Layer 4\cite{OSI} and above protocols, we considered the widely accepted combination of CoAP\cite{rfc7252} over DTLS\cite{rfc6347}(optional) over UDP\cite{rfc768}\footnote{CoAPs is equivalent to CoAP over DTLS.}. \Cref{Protocols} summarises our choice of protocol stack.

\begin{table}[!h]
	\centering
	\input{tbl/protocols.tex}
	\caption{Protocol stack for our experiments(* is optinal)\label{Protocols}}
\end{table}

@@

%Avaible security options.
For our settings, there are two schemes available for packet encryption, namely 802.15.4 Security\cite{802154} and DTLS\cite{rfc6347}. 802.15.4 Security is provided by noncoresec\cite{noncoresec} which implements 802.15.4 authenticated encryption with AES-128 CCM*\cite{CCM} using a hard-coded key shared by the whole 6LoWPAN network. tinyDTLS\cite{tinydtls} provides a minimum DTLS implementation that supports only two ciphersuites which are TLS\_PSK\_WITH\_AES\_128\_CCM\_8\cite{rfc6655} and TLS\_ECDHE\_ECDSA\_WITH\_AES\_128\_CCM\_8\cite{rfc6655} respectively. We do not concern the case where both 802.15.4 Security and DTLS are enabled simultaneously as the resulting packets exceeded the 802.15.4 MTU\footnote{Maximum Transmission Unit.} and thus not practical.

>>>>

\section{A Typical IoT Protocol Stack\label{sec: IotProtocols}}

%IoT Protocols
There are large number of protocols, which have been proposed for different IoT applications adapting to various requirements. For example, some smart houses simply use WiFi for connectivity and VANETs\footnote{Vehicular ad hoc networks} may adopt DSRC\cite{DSRC}.

%WSN Protocols
In this paper we focus on 6LoWPAN\cite{rfc4944} which is based on the 802.15.4\cite{802154} standard. These standards are designated for constrained environments such as Wireless Sensor Networks,  but other competing standards exist at different layers. Bluetooth Low Energy(BLE)\cite{BLE} is a strong competitor to 802.15.4 as well as the LiFi\cite{LiFi} technology. Zigbee\cite{Zigbee} was originally intended as a collective protocol over 802.15.4 but it has been recently adapted to IP-based network in ZigbeeIP\cite{ZigbeeIp}. The RIME stack\cite{RIME} proposes a set of non-layered primitives over 802.15.4 but it is likely to be phased-out due to the lack of of interoperability with the TCP/IP protocol stack.

%6LoWPAN OS
6LoWPAN thus is the most popular standard for low power networks, and thus it is supported by several competing IoT Operating Systems, including Contiki OS\cite{Contiki}, OpenWSN\cite{OpenWSN}, FreeRTOS\cite{FreeRTOS} and the recent RIOT\cite{RIOT}. We chose Contiki OS for our experiments because it is easy to customise.

\subsection{Our experimental network}
%Say something about devices, and choice of OS, etc.
%Hardware
Our experimental network is constructed using two different devices. These are a TelosB\cite{TelosB} and a CC2538\cite{CC2538}. The TelosB is a low cost sensor powered by an MSP430 with an AES co-processor. It represents typical low-end devices. The CC2538 is the high end device powered by an ARM Cortex-M3 with multiple cryptographic processors including AES, RSA, SHA-2 and ECC, suggesting that it is suitable to develop secure applications. %Should I cite our previous paper...?

%Software
Both devices are supported by the Contiki OS. We adopted the default settings of the Contiki OS, except for enabling 802.15.4 security\cite{802154} for some experiments. Note that the Contiki MAC\cite{ContikiMAC} is chosen by default over TSCH\cite{TSCH}. For Layer 4\cite{OSI} and above protocols, we went with the widely accepted combination of CoAP\cite{rfc7252}, and DTLS\cite{rfc6347}(optional) over UDP\cite{rfc768}\footnote{CoAPs is equivalent to CoAP over DTLS.}. \Cref{Protocols} summarises our choice of protocol stack.

\begin{table}[!h]
	\centering
	\begin{tabular}{|c|c|}
		\hline
		Physical                        & \multirow{2}{*}{802.15.4} \\ \cline{1-1}
		Link                          &                           \\ \hline
		Network                       & 6LoWPAN                   \\ \hline
		\multirow{2}{*}{Transmission} & UDP                       \\ \cline{2-2} 
		& DTLS*                     \\ \hline
		Application                   & CoAP / CoAPs*             \\ \hline
	\end{tabular}

	\caption{Protocol stack for our experiments(* is optinal)\label{Protocols}}
\end{table}

@@

%Avaible security options.
\subsubsection{802.15.4 and DTLS}
In our setting, there are two standards available for packet encryption, namely 802.15.4 security\cite{802154} and DTLS\cite{rfc6347}. 802.15.4 security is provided by the noncoresec\cite{noncoresec} API, which implements 802.15.4 authenticated encryption with AES-128 CCM*\cite{CCM} using a hard-coded key shared by the whole 6LoWPAN network. We chose tinyDTLS\cite{tinydtls}  as library for the DTLS protocols, because it provides a minimum DTLS implementation that supports two ciphersuites which are TLS\_PSK\_WITH\_AES\_128\_CCM\_8\cite{rfc6655} and TLS\_ECDHE\_ECDSA\_WITH\_AES\_128\_CCM\_8\cite{rfc6655} respectively. Evidently, they both utilise AES-128 CCM* as the packet encryption method.
